---
title: "The Correlation between Attitudes towards Immigrants and Voting Behavior"
author: Yun-Tsz Tsai
date-format: "MMMM DD, YYYY"
format: pdf
---
# Introduction
It is not unheard that one's party preference are highly associated with one's opinions on certain issues. Parties that are on the right end of the spectrum are more likely to be anti-immigrants while parties that are on the left end of the spectrum tend to have more open policy on immigrants. As a result, it will be reasonably to assume that if an individual has negative view on immigrants is more likely to support a party whose ideology aligns with this and vice versa.

In this study, I will investigate the correlation between individuals' views on immigrants and their voting behavior with a focus on Finnish population. The hypothesis is formulated as follows,

*Do peopel who have more negative views on immigrants more likely to vote for praties that are right-leaning?*

The data being used in this study is from round 11 (2023/24) of European Social Suvery (ESS). 

link to find the list of variables in this data set <https://ess.sikt.no/en/datafile/242aaa39-3bbb-40f5-98bf-bfb1ce53d8ef>

```{r}
#| echo: false
#| message: false
#| warning: false
library(tidyverse)
library(haven)

# load the data
ess2023_fi <- read_dta('./ESS11_fi.dta')

# independent v = 'prtvtffi' which stands for party voted for in last election
variables <- c('prtvtffi', 'imbgeco', 'gndr', 'edlvdfi', 'agea') # dtype = double
ess2023_fi <- ess2023_fi %>% 
                select(variables) %>% 
                filter(complete.cases(.)) 


# remove missing values and convert some columns to factors
ess2023_fi <- ess2023_fi %>% 
  mutate(party = zap_missing(prtvtffi),
         party = as_factor(party),
         party = fct_lump_n(party, 3),
         imbgeco = as.integer(imbgeco),
         gndr = as_factor(gndr),
         edlvdfi = zap_missing(edlvdfi),
         # highly educated or not
         edlvdfi = as_factor(ifelse(edlvdfi >= 9, 1, 0)), 
         agea = as.integer(agea)) %>% 
         filter(gndr != 'No answer')

# relevel and set National Coalition Party as a reference category
ess2023_fi <- within(ess2023_fi, party <- relevel(
  party, ref = 'The National Coalition Party'))

```

::: {.landscape}
```{r}
#| echo: false
#| message: false
#| warning: false
#| tbl-cap: 'Descriptive Statistics on Voter Arributes'
#| label: tbl-descriptive
library(gtsummary)
library(modelsummary)
library(kableExtra)
# descriptive data on the sample
ess2023_fi %>% 
  mutate(group = case_when(
    party ==  'The National Coalition Party' ~ 1,
    party == 'Social Democratic Party' ~ 2,
    party == 'True Finns' ~ 3,
    party == 'Other' ~ 4
  )) %>% 
  mutate(group = factor(group, labels = c('NCP', 'SDP', 'TF', 'Other'))) %>% 
  select(group, 
         Sex = gndr,
         Age = agea,
         `Higher Education` = edlvdfi,
         `Opinions on Immigrants` = imbgeco) %>%
  tbl_summary(by = group, include = c(Sex, Age, 
                                      `Higher Education`, `Opinions on Immigrants`)) %>% 
  add_overall() %>% 
  remove_row_type(variables = Sex,
                  type = 'level',
                  level_value = 'No answer') %>% 
   modify_footnote_header("Source: European Social Survey Round 11,
                                 Higher values in *Opinion on Immigrants*
                                 suggest more positive views.",
                          , columns = all_stat_cols())

```
:::

::: {.landscape}
```{r}
#| echo: false
#| message: false
#| warning: false
#| tbl-cap: 'Odds Ratios from Logistic Regression Models of Party Choice'
#| label: md-statistics
library(nnet)
library(broom)
library(modelr)
library(modelsummary)

# estimation using multinomial logistic regression models
 # model 1 with opinions on immigrants as the only idependent variable
mod1 <- multinom(party ~ imbgeco, data = ess2023_fi, trace = FALSE)

 # model 2 with opinons + demographic attributes
mod2 <- multinom(party ~ imbgeco + gndr + edlvdfi + agea, data = ess2023_fi,
                 trace = FALSE)
# making a table
models <- list(mod1, mod2)
modelsummary(models, shape =  term ~ model + response,
             fmt = 2,
             statistic = 'conf.int',
             exponentiate = TRUE,
             gof_omit = 'RMSE')
```
:::

```{r}
#| echo: false
#| message: false
#| warning: false
#| tbl-cap: 'Goodness of Model Fit'
#| label: md-fit
library(lmtest)
library(DescTools)
library(gt)
# model fit statistics
lktest <- lrtest(mod1, mod2)
# Negelkerke's pseuso's R^2
pse_R1 <- round(PseudoR2(mod1, which = 'Nagelkerke'), 2)
pse_R2 <- round(PseudoR2(mod2, which = 'Nagelkerke'), 2)
# share of correct prediction
 # model 1
pred_mod1 <- predict(mod1, type = 'class')
sc_mod1 <- (mean(ess2023_fi$party == pred_mod1)) *100
 # model 2
pred_mod2 <- predict(mod2, type = 'class')
sc_mod2 <- (mean(ess2023_fi$party == pred_mod2)) *100

# put the information into a table
results <- tibble('term' = c('Likelihood Ratio Test', 
                      'Pseuso R²', 
                      'Share of Correct Prediction (%)'),
                  'Model 1' = c(lktest$Chisq[1], pse_R1, sc_mod1),
                  'Model 2' = c(lktest$Chisq[2], pse_R2, sc_mod2))
results %>% 
  gt() %>% 
  fmt_number(
    columns = everything(),
    decimal = 1
  ) %>% 
  cols_label(
    term = 'Metrics',
    `Model 1` = 'Model 1',
    `Model 2` = 'Model 2')
```

